---
title: "EDS 221 Day 4 Tasks and Activities"
author: "Halina Do-Linh"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Task 1

```{r}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
transects <- c("Transect A", "Transect B", "Transect C")

wk_trans_matrix <- matrix(ncol = length(transects), nrow = length(weekdays))

for (i in seq_along(transects)) {
  for (j in seq_along(weekdays)) {
    wk_trans_fill <- print(paste(weekdays[j], "-", transects[i]))
    wk_trans_matrix[j, i] <- wk_trans_fill
  }
}

wk_trans_matrix

# j is row, i is column
# inner loop rows
# outer loop columns
```

## Task 2 

```{r}
force <- function(mass, acceleration) {
  force = mass * acceleration
  return(paste("The resulting force is", force, "Newtons"))
}

force(12, 3)
```

## Task 3
Write a function called name_column that, given a data frame name and a number, will return the column name associated with that column number. Add an error message if the column number is < 1, or if the number exceeds the number of columns in the data frame. Test the function using the mtcars data frame.

```{r}
name_column <- function(df, number) {
  return(colnames(df[number]))
 
   if (number < 1) {
    stop("Number cannot be less than 1")
  } else if (number > ncol(df)) {
    stop("Number cannot be more than number of columns in data frame")
  }
}

name_column(mtcars, 15)

# need to come back to this
```


## Task 4

 W=aLb

```{r}
# alternatively I could make vectors for each column and use data.frame() with those vectors

fish_parms <- data.frame(sci_name = c("Chanos chanos", "Sphyraena barracuda", "Caranx ignobilis"), common_name = c("Milkfish", "Great barracuda", "Giant trevally"), a_est = c(0.0905, 0.0181, 0.0353), b_est = c(2.52, 3.27, 3.05))

fish_weight <- function(fish_name, tot_length) {
  select_fish <- dplyr::filter(fish_parms, common_name == fish_name)
  a = select_fish$a_est
  b = select_fish$b_est
  fish_weight = a * tot_length ^ b
  return(fish_weight)
}

fish_weight("Milkfish", 6)

```


## Task 5

```{r}
wave_power <- function(wave_height, period) {
  wave_power = 0.5 * (wave_height)^2 * period
  return(wave_power)
}

wh_seq <- seq(from = 0, to = 3, by = 0.2)

wave_power(wh_seq, 8)

```

## Task 6 

```{r}


wave_power <- function(water_depth, wave_height, period) {
  if (water_depth > 12) {
    wave_power = 0.5 * (wave_height)^2 * period
    message("Deep water equation used")
    return(wave_power)
  }
  if (water_depth <= 12) {
    wave_power = 0.81 * wave_height ^ 2 * period
    message("Shallow water equation used") 
    return(wave_power)
  }
 
}

wave_power(5, 12, 12)
```

